---
title: "My Travel Map"
author: Daniel Jang
date: "`r format(Sys.time(), '%B %d, %Y')`"
date-format: iso
format: 
  html: 
    embed-resources: true
execute:
  echo: false
  warning: false
  error: false
  message: false
editor: source
---

```{r}
library(tidyverse)
library(tidyselect)
library(tibble)
library(sf)
library(tigris)
library(lubridate)
library(tidygeocoder)

library(gt)
library(gtsummary) # nice count visualizations
library(forcats) # a specific data wrangling issue
```

# The United States

After my semester abroad, I made a goal of travelling to every U.S. state. I am documenting my progress here because I can make nice visualizations and data

```{r}
# The dataset: initialization from an example list of trips I made
data.frame(
  type             = c("Leisure", "Leisure"),
  desc             = c("Friend", "Friend"), # can just be text
  state            = c("DC", "Texas"),
  cities           = c("Washington", "Austin"), #comma-separated 
  start_place      = c("New York City, NY", "Boston, MA"),
  start_date       = ymd(c("2025-07-12", "2025-08-15")),
  end_place        = c("New York City, NY", "Boston, MA"),
  end_date         = ymd(c("2025-07-12", "2025-08-18")),
  transport_method = c("Train", "Plane"),
  redeye_indicator = c("Yes", "No"))
```

```{r}
add_trip <- function(file = "trips.rds") {

  # Load existing trips
  trips <- readRDS(file)

  
  cat("\n--- New Trip Entry ---\n")
  type <- readline("Type (Leisure/Business): ")
  desc <- readline("Description: ")

  
  state <- readline("State (Full name): ")
  cities <- readline(
    "Cities visited (comma-separated): "
  )


  start_place <- readline("Start location (City, State): ")
  start_date <- ymd(
    readline("Start date (YYYY-MM-DD): ")
  )

  
  end_place <- readline("End location (City, State): ")
  end_date <- ymd(
    readline("End date (YYYY-MM-DD): ")
  )

  
  transport_method <- readline(
    "Transport method (e.g. Plane, Train, Car): "
  )
  redeye_indicator <- readline(
    "Redeye? (Yes/No): "
  )

  new_trip <- tibble(
    type = type,
    desc = desc,
    state = state,
    cities = cities,
    start_place = start_place,
    start_date = start_date,
    end_place = end_place,
    end_date = end_date,
    transport_method = transport_method,
    redeye_indicator = redeye_indicator
  )

  # Append
  trips <- bind_rows(
    trips,
    new_trip
  )

  # Save automatically
  saveRDS(
    trips,
    file
  )

  cat("\nTrip saved to", file, "\n")

  invisible(trips)
}
```

Run this code to sort everything

```{r}
#| eval: false

temp <- readRDS("trips.rds")

# example code to rewrite mistakes
fix(temp)


# rearrange and make sure everything is as it should be
temp <- temp |>
  dplyr::arrange(desc(start_date)) |>
  mutate(start_date = as.Date(start_date),
         end_date = as.Date(end_date))

#ideas: solo trip indicator, what company did i use to travel
saveRDS(temp, "trips.rds")
```

*NEW TRIP ENTRY*

```{r}
#| echo: true
#| eval: false
add_trip()
```

> Need to fix this so that it's not a whole list printed in the html file

```{r}
daniel_travel_map <- readRDS("trips.rds")

daniel_travel_map |> 
  dplyr::select(-desc, -start_date, -end_date) |>
  knitr::kable(format = "html")
```

This does not count the states I have actually lived in. Not all trips I have done in my life are included. These are just ones I thought were memorable. I also don't count just passing through a state. 

```{r}
visit_counts <- daniel_travel_map |>
  count(state, name = "visits")

# US states shapefile
us <- states(cb = TRUE) |>
  shift_geometry() #for alaska and hawaii

# Join counts to map
map_data <- us |>
  left_join(
    visit_counts,
    by = c("NAME" = "state")
  ) |>
  mutate(
    visits = ifelse(is.na(visits), 0, visits)
  )

st_bbox(map_data)

# Plot
ggplot(map_data) +
  geom_sf(aes(fill = visits)) +
  coord_sf(xlim = c(-3500000, 3500000), ylim = c(-2500000, 2500000), expand = FALSE) + 
  scale_fill_viridis_c(name = "Trips") +
  scale_fill_gradient(low = "white", high = "blue") + 
  theme_minimal() 
```

Make a shiny app where I have a date slider. I want to see the timeline on when I visit states
Or maybe something that hovers so I can see where and when I last went

Also a visual that has lines between states to show a rough path towards everything.